---
title: "p8105_hw2_mm5354"
author: "Mengran Ma"
date: "2018/9/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#Problem 1
#####Read and clean the data
```{r}
transit_data = 
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names()
```
#####Retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 
```{r}
select(transit_data, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)
```
#####Convert the entry variable from character (YES vs NO) to a logical variable 
```{r}
recode(transit_data$entry, YES = 1, NO = 0)
```

#####Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?
```{r}
nrow(transit_data)
ncol(transit_data)
```

#####How many distinct stations are there?
```{r}
distinct(transit_data, station_name)
```
#####What proportion of station entrances / exits without vending allow entrance?
```{r}
sum(transit_data$vending == "NO")
sum(transit_data$vending == "NO" & transit_data$entry == "YES")/sum(transit_data$vending == "NO", TRUE)
```
####Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?


#Problem 2
#####Read and clean the Mr. Trash Wheel sheet:
#####1. specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
#####2. use reasonable variable names
#####3. omit rows that do not include dumpster-specific data
#####4. rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r}
P2 = 
  readxl::read_excel("./HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 1)
P2 = janitor::clean_names(P2)
```
#####Omit rows that do not include dumpster-specific data
```{r}
P2[!is.na(P2$Dumpster), ]
```

#####Rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r}
P2SportsBalls = ceiling(P2$sports_balls)
as.integer(P2SportsBalls)
```
####Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r}
P22016 = 
  readxl::read_excel("./HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, skip = 1)
P22016 = janitor::clean_names(P22016)
```
```{r}
P22017 = 
  readxl::read_excel("./HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, skip = 1)
P22017 = janitor::clean_names(P22017)
```

```{r}
P22016noNa = na.omit(P22016, cols = seq_along(P22016))
NewP22016 = mutate(P22016noNa, year = "2016")
P22017noNa = na.omit(P22017, cols = seq_along(P22017))
NewP22017 = mutate(P22017noNa, year = "2017")

```
```{r}
Join_data = left_join(NewP22016, NewP22017, by = "month")
Join_data[, 1] <- month.name[sapply(Join_data[, 1], as.integer)]
Join_data
```
For available data,the total precipitation in 2017 is `r sum(NewP22017$total)` What was the median number of sports balls in a dumpster in 2016?



```{r}

```

