---
title: "p8105_hw2_mm5354"
author: "Mengran Ma"
date: "2018/9/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(httr)
library(jsonlite)
```

#Problem 1
#####Read and clean the data
```{r}
transit_data = 
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names()
```
#####Retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 
```{r}
select(transit_data, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)
```
#####Convert the entry variable from character (YES vs NO) to a logical variable 
```{r}
recode(transit_data$entry, YES = TRUE, NO = FALSE)
```

#####Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?
```{r}
nrow(transit_data)
ncol(transit_data)
```

#####How many distinct stations are there?
```{r}
distinct(transit_data, station_name)
```
#####What proportion of station entrances / exits without vending allow entrance?
```{r}
sum(transit_data$vending == "NO")
sum(transit_data$vending == "NO" & transit_data$entry == "YES")/sum(transit_data$vending == "NO", TRUE)
```
####Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?


#Problem 2
#####Read and clean the Mr. Trash Wheel sheet:
#####1. specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
#####2. use reasonable variable names
#####3. omit rows that do not include dumpster-specific data
#####4. rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r}
P2 = 
  readxl::read_excel("./HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 1)
P2 = janitor::clean_names(P2)
```
#####Omit rows that do not include dumpster-specific data
```{r}
P2[!is.na(P2$Dumpster), ]
```

#####Rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r}
P2SportsBalls = ceiling(P2$sports_balls)
as.integer(P2SportsBalls)
```
####Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r}
P22016 = 
  readxl::read_excel("./HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, skip = 1)
P22016 = janitor::clean_names(P22016)
```
```{r}
P22017 = 
  readxl::read_excel("./HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, skip = 1)
P22017 = janitor::clean_names(P22017)
```

```{r}
P22016noNa = na.omit(P22016, cols = seq_along(P22016))
NewP22016 = mutate(P22016noNa, year = "2016")
P22017noNa = na.omit(P22017, cols = seq_along(P22017))
NewP22017 = mutate(P22017noNa, year = "2017")

```
```{r}
Join_data = left_join(NewP22016, NewP22017, by = "month")
Join_data[, 1] <- month.name[sapply(Join_data[, 1], as.integer)]
Join_data
```
Number of observations in both resulting datasets: `r nrow(P2)` for Mr. Trash Wheel sheet and `r nrow(Join_data)` for precipitation data for 2016 and 2017 in total. For available data,the total precipitation in 2017 is `r sum(NewP22017$total)` The median number of sports balls in a dumpster in 2016 is `r median(na.omit(cbind(P2$sports_balls[P2$year == 2016])))`.

#Problem 3
For this question:

1. format the data to use appropriate variable names;
2. focus on the “Overall Health” topic
3. exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation
4. structure data so that responses (excellent to poor) are variables taking the value of Data_value
5. create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

6. How many unique locations are included in the dataset? Is every state represented? What state is observed the most?
7. In 2002, what is the median of the “Excellent” response value?
8. Make a histogram of “Excellent” response values in the year 2002.
9. Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010
```{r}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets", force = TRUE)

library(p8105.datasets)
data("brfss_smart2010")
P3 = janitor::clean_names(brfss_smart2010)
P3
```

